### 2018-04-01 paper/sexSelectionSpacing-ver1.tex
Edit High Education Women

Women with eight or more years of education appear the least likely
to follow the pattern of shorter spacing when they have only girls
compared to only boys before sex selection became available as 
Table \ref{tab:median_sex_ratio_high} shows.
Part of the problem may be that there are few women with that level
of education for the 1972--1984 period.
Hence, only the results for the first and second spells in 
both urban and rural areas are likely to be reliable.
Furthermore, there are relatively few women with three boys who
go on to having a fourth birth, both in urban and rural areas.

Birth spacing increases substantially over time, 
especially for urban women. 
The median duration for urban women increase between a quarter
and three-quarters of a year, whereas the increase for rural women 
is around a quarter of a year.
The spacing between marriage and the first birth is still the
exception, and for this group of women, there is a decline of
about a month for both urban and rural areas.

In urban areas, the increase in spacing with only girls is so 
substantial that women with only girls have longer spacing to the 
next birth than any of the other sex compositions from 1985 onward.%
\footnote{
The exception is the second spell for the 1985--1994 period, where women
with one girl have 0.4 months longer spacing the women with one boy.
}
This reversal in the standard spacing pattern is substantial, 
especially for the third and fourth spells.
For the latest period, the difference between only girls than the
other sex compositions is between 3.4 and 5.4 months for the third
and fourth spell.
Furthermore, the differences are statistically significant, except
for only boys for the fourth spell because few women have a fourth
birth if they have three boys already.

@@

When trying to understand the strength of son preference, it is interesting
that the sex ratio is also statistically significantly different
from the natural rate in the case where women already have one
son for the third and fourth spells.
Again, the sex ratio in the presence of one son for the third and
fourth spells are higher in urban areas than rural areas, although
the difference is less than for women with only girls.
Hence, it is possible that women are still willing to use sex selection
even after giving birth to one. 
Alternatively, this behavior may be in response to either experienced
or expected mortality of the first son born.

@@

Both of the last two results are different from prior studies using NFHS data.
There are two potential explanations.
First, the NFHS-4 is substantially larger than the three prior surveys.
Hence, it is possible that the effects have been there all along, but 
we did not have the power to detect them.
Second, it is possible that, at least, the first-born results are an
artifact of any recall error not captured by the data restrictions.
Certain types of recall error would also explain why the sex ratio for
the second spell and one boy is statistically significantly \emph{below} 
the natural sex ratio \citep{Portner2018b}.

>>>>

Women with eight or more year of education mostly followed 
the traditional spacing pattern before sex selection became 
available as shown in Table \ref{tab:median_sex_ratio_high}.
Compared to the two other education groups there are,
however, more situations where the pattern is less clear.
Part of the problem may be that there are few women in
this group for the 1972--1984 period, and, hence, only 
the results for the first and second spells in both urban 
and rural areas are likely reliable.

Birth spacing increases substantially over time, 
especially for urban women. 
The median duration for urban women increases between a half
and a full year, whereas the increase for rural women 
is between a quarter and half a year.
The spacing between marriage and the first birth is still the
exception, and for this group of women, there is a decline of
about a month for both urban and rural areas.

In urban areas, the increase in spacing with only girls is so 
substantial that women with only girls have longer spacing to the 
next birth than any of the other sex compositions from 1985 onward.%
\footnote{
The exception is the second spell for the 1985--1994 period, where women
with one girl have 0.1 months shorter spacing the women with one boy.
}
This reversal in the standard spacing pattern is substantial, 
especially for the third and fourth spells.
For the latest period, the difference between only girls than the
other sex compositions is between 4.8 and 9.1 months for the third
and fourth spell.
Furthermore, the differences are statistically significant, except
for only boys for the fourth spell because few women have a fourth
birth if they have three boys already.

@@

When trying to understand the strength of son preference, it is interesting
that the sex ratio is also statistically significantly different
from the natural rate in the case where women already have one
son for the third and fourth spells.
Again, the sex ratio in the presence of one son for the third and
fourth spells are higher in urban areas than rural areas, although
the difference is less than for women with only girls.
Hence, it is possible that women are still willing to use sex selection
even after giving birth to one, although this behavior may also 
be in response to either experienced
or expected mortality of the first son born.

@@

Both of the last two results are different from prior studies using NFHS data.
There are two potential explanations.
First, the NFHS-4 is substantially larger than the three prior surveys.
Hence, it is possible that the effects have been there all along, but 
we did not have the power to detect them.
Second, it is possible that at least the first-born results are an
artifact of any recall error not captured by the data restrictions.
Certain types of recall error would also explain why the sex ratio for
the second spell and one boy is statistically significantly \emph{below} 
the natural sex ratio \citep{Portner2018b}.
